\SW: Crowd-sourcing the Discovery of Gravitational Lenses

### 2014-06-02 doc/sw-system.tex
New numbers, from re-run of SWAP to 'correct' end date

We define the following properties of the crowd, as characterised by their
agents, and plot the distributions of their logarithms in
\Fref{fig:crowd:cornerplot}. 
%
\begin{description}
%
\item{\noindent\bf ``Effort:''} The number of test images, $\effort$, classified by a
volunteer. In stage 1, the mean effort per agent was 259; in the shorter stage
2 it was 81.
%
\item{\noindent\bf ``Experience:''} The number of training images, $\experience$,
classified by a volunteer. In stage 1, the mean experience per agent was 29;
in stage 2 (where the training image fequency was set higher) it was 34.
%
\item{\noindent\bf ``Skill:''} The expectation value of the information gain, $\skill$
should  the next subject classified have lens probability 0.5
\Aref{appendix:swap},  in bits. Random classifiers have $\skill = 0.0$,
perfect classifiers have $\skill = 1.0$. All agents start with $\skill = 0.0$;
this increases as training subjects are classified, and the agent's estimates
of its confusion matrix elements improve. In stage 1, the mean skill per agent
was 0.04 bits; in  stage 2 it was 0.05.
%
\item{\noindent\bf ``Contribution:''} The integrated skill over a volunteer's test
subject classification history, and represents the total  contribution to the
project that volunteer (see the appendix for more discussion of this
quantity). In stage 1, the mean contribution per agent was 34.3 bits; in 
stage 2 it was 33.5.
%
\item{\noindent\bf ``Information:''} The total information $\information$ generated by
the  agent during the volunteer's classification activity. This quantity
depends on the value of each subject's lens probability when that subject was 
presented to the volunteer (\Aref{appendix:swap}), and so there is an element
of luck involved with this quantity: if you never see a high probability
subject, it's hard to generate a large amount of information. You make your
own luck by classifying more subjects.
%
\end{description}

@@

\Fref{fig:crowd:cornerplot} shows the \sw crowd to have quite broad
distributions of logarithmic effort, skill, and contribution. To better
quantify the contributions made by the volunteers, we show their cumulative
distribution on a linear scale in \Fref{fig:crowd:cumulplot}. This plot shows
clearly the importance of the hardest-working, most active volunteers: at
stage 1,  1.0\% of the volunteers -- 374 people -- made 90\% of the
contribution.  At stage 2, where it was not possible to make as
many classifications before running out of subjects, 7.2\% of the volunteers
-- 141 people -- made 90\% of the contribution.

However, it is not the case that only these small groups were capable of
making this large contribution: 9119 volunteers were ``experienced'' in that
they had all classified at least 10 training images; the 374 highest
contributing volunteers make up just 4\% of this experienced volunteer pool.
The cumulative distribution of agent skill is shown in
\Fref{fig:crowd:cumulskillplot}: these distributions are significantly broader
than the corresponding distributions of agent contribution in 
\Fref{fig:crowd:cumulplot}. The most skilled 20\% of agents possess only  79\%
of the skill at stage 1, and 77\% at stage 2. The inexperienced volunteers
also possess a significant fraction of the skill: the most skillful 20\% of
experienced volunteers (1824 people)  possess just 43\% of the total skill.
The level of contribution made at \sw by experienced volunteers is largely a
matter of choice (or perhaps, availability of time!).

@@

We now look at the transition between stage 1 and stage 2 in more detail. Of
the 1964 volunteers who took part in the stage 2 classification round, only
783 were veterans from stage 1. However, these volunteers showed significantly
higher skill levels than the new users, and also made more classifications
(and hence made greater contributions. This is illustrated in  
\Fref{fig:crowd:stage1vstage2}. In this figure, point size is proportional to
information generated: the higher skill, higher effort stage 1 veterans
generated most of the information.

@@

\Tref{tab:crowd:contributions} shows the total effort, contribution, skill and
information generated in both stage 1 and 2 of the CFHTLS project, with the
total numbers of agents and subjects for comparison. These numbers allow us to
quantify the efficiency of the system. The contribution per classification is
defined in terms of a hypothetical subject with lens probability of 0.5; one
bit of information is needed to update such a subject's lens probability to
either zero or one. This means that a maximally complete classification stage
would yield a total contribution (summed over all agents) equal to the number
of subjects. The ratio of this hypthetical optimum to the actual total
contribution is therefore a measure of the stage's inefficiency. We find our
inefficiency (by dividing column 2 by column 3) to be 33\% and 17\% in stage 1
and 2 respectively. In stage 1, this inefficiency is due to the daily
processing: we were not able to retire subjects fast enough, and so they,
remained in the system, being over-classified. Only 3644943 classifications
were needed to retire all the subjects: the ratio of this to the total number
made is 34\%. (The remaining 1\% is due to not all subjects being classified
to 1 or 0 probability.) At stage 2, we did not retire any subjects at all; the
inefficiency in this case was by design, to give everyone a chance to
appreciate what they had found together!

The total skill of the crowd, computed by summing the skill of all the agents,
is, similarly, a measure of the effective crowd size: a crowd of perfect
classifiers would be this size. The stage 1 crowd was equivalent to a team of
1472.1 perfect classifiers; the stage 2 crowd was equivalent to a team of
102.4 perfect classifiers.

The total information generated during the survey is harder to interpret,
because it's exactly the high probability candidates that we leave in the
system that give rise to most of the information generated. What we can do is
divide the total information generated by the amount of information it takes
to classify a \sw subject all the way to the detection threshold (lens
probability 0.95), divide by this number, and then multiply by the survey
inefficiency to get a very rough estimate for the effective number of
detections corresponding to the crowd's contribution. With a prior probability
of $2\times10^{-4}$, it takes 11.6 bits to make a detection; the effective
number of detections contributed by the stage 1 and stage 2 volunteers is then
1736 and 24 respectively. These figures lie roughly between the numbers of
detections and non-rejections given in column 7 of
\Tref{tab:crowd:contributions}. The uncertainty in the interpretation of the
information generated provides further justification for our focus on the
expected information gain as a measure of volunteer contribution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}
\begin{center}
\caption{Total crowd and subject sample properties from the CFHTLS project.}
\label{tab:crowd:contributions}
\begin{tabular}{cccccccc}
  \hline
  \hline {Stage} & Subjects & Contribution                          & Agents & Skill      & Classifications          & Candidates & Information \\
                 & $\Ns$    & $\sum_k^{\Nv} \contribution_k$ (bits) & $\Nv$  & $\sum_k^{\Nv} \skill_k$ (bits) & $\sum_k^{\Nv} \thiseffort$ & $\Ncands$  & $\sum_j^{\Ns}\sum_k^{\Nv} \information_{j,k}$ (bits) \\
  \hline 
            1    & $423683$ & $1269574.7$ & $36985$ & $1472.1$ & $10654285$ & $43 (3225)$ & $61032.7$ \\
            2    & $3679$   &   $21895.8$ &  $1964$ &  $102.4$ &   $224745$ & $91 (1391)$ &  $1641.9$ \\
  \hline \hline
\end{tabular}
\medskip\\
\end{center}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

We define the following properties of the crowd, as characterised by their
agents, and plot the distributions of their logarithms in
\Fref{fig:crowd:cornerplot}. 
%
\begin{description}
%
\item{\noindent\bf ``Effort:''} The number of test images, $\effort$, classified by a
volunteer. In stage 1, the mean effort per agent was 263; in the shorter stage
2 it was 81.
%
\item{\noindent\bf ``Experience:''} The number of training images, $\experience$,
classified by a volunteer. In stage 1, the mean experience per agent was 29;
in stage 2 (where the training image fequency was set higher) it was 34.
%
\item{\noindent\bf ``Skill:''} The expectation value of the information gain, $\skill$
should  the next subject classified have lens probability 0.5
\Aref{appendix:swap},  in bits. Random classifiers have $\skill = 0.0$,
perfect classifiers have $\skill = 1.0$. All agents start with $\skill = 0.0$;
this increases as training subjects are classified, and the agent's estimates
of its confusion matrix elements improve. In stage 1, the mean skill per agent
was 0.04 bits; in  stage 2 it was 0.05.
%
\item{\noindent\bf ``Contribution:''} The integrated skill over a volunteer's test
subject classification history, and represents the total  contribution to the
project that volunteer (see the appendix for more discussion of this
quantity). In stage 1, the mean contribution per agent was 34.3 bits; in 
stage 2 it was 33.5.
%
\item{\noindent\bf ``Information:''} The total information $\information$ generated by
the  agent during the volunteer's classification activity. This quantity
depends on the value of each subject's lens probability when that subject was 
presented to the volunteer (\Aref{appendix:swap}), and so there is an element
of luck involved with this quantity: if you never see a high probability
subject, it's hard to generate a large amount of information. You make your
own luck by classifying more subjects.
%
\end{description}

@@

\Fref{fig:crowd:cornerplot} shows the \sw crowd to have quite broad
distributions of logarithmic effort, skill, and contribution. To better
quantify the contributions made by the volunteers, we show their cumulative
distribution on a linear scale in \Fref{fig:crowd:cumulplot}. This plot shows
clearly the importance of the hardest-working, most active volunteers: at
stage 1,  1.0\% of the volunteers -- 375 people -- made 90\% of the
contribution.  At stage 2, where it was not possible to make as
many classifications before running out of subjects, 7.2\% of the volunteers
-- 141 people -- made 90\% of the contribution.

However, it is not the case that only these small groups were capable of
making this large contribution: 9097 volunteers were ``experienced'' in that
they had all classified at least 10 training images; the 375 highest
contributing volunteers make up just 4\% of this experienced volunteer pool.
The cumulative distribution of agent skill is shown in
\Fref{fig:crowd:cumulskillplot}: these distributions are significantly broader
than the corresponding distributions of agent contribution in 
\Fref{fig:crowd:cumulplot}. The most skilled 20\% of agents possess only  79\%
of the skill at stage 1, and 77\% at stage 2. The inexperienced volunteers
also possess a significant fraction of the skill: the most skillful 20\% of
experienced volunteers (1820 people)  possess just 43\% of the total skill.
The level of contribution made at \sw by experienced volunteers is largely a
matter of choice (or perhaps, availability of time!).

@@

We now look at the transition between stage 1 and stage 2 in more detail. Of
the 1964 volunteers who took part in the stage 2 classification round, only
774 were veterans from stage 1. However, these volunteers showed significantly
higher skill levels than the new users, and also made more classifications
(and hence made greater contributions. This is illustrated in  
\Fref{fig:crowd:stage1vstage2}. In this figure, point size is proportional to
information generated: the higher skill, higher effort stage 1 veterans
generated most of the information.

@@

\Tref{tab:crowd:contributions} shows the total effort, contribution, skill and
information generated in both stage 1 and 2 of the CFHTLS project, with the
total numbers of agents and subjects for comparison. These numbers allow us to
quantify the efficiency of the system. The contribution per classification is
defined in terms of a hypothetical subject with lens probability of 0.5; one
bit of information is needed to update such a subject's lens probability to
either zero or one. This means that a maximally complete classification stage
would yield a total contribution (summed over all agents) equal to the number
of subjects. The ratio of this hypthetical optimum to the actual total
contribution is therefore a measure of the stage's inefficiency. We find our
inefficiency (by dividing column 2 by column 3) to be 33\% and 17\% in stage 1
and 2 respectively. In stage 1, this inefficiency is due to the daily
processing: we were not able to retire subjects fast enough, and so they,
remained in the system, being over-classified. Only 3705704 classifications
were needed to retire all the subjects: the ratio of this to the total number
made is 34\%. (The remaining 1\% is due to not all subjects being classified
to 1 or 0 probability.) At stage 2, we did not retire any subjects at all; the
inefficiency in this case was by design, to give everyone a chance to
appreciate what they had found together!

The total skill of the crowd, computed by summing the skill of all the agents,
is, similarly, a measure of the effective crowd size: a crowd of perfect
classifiers would be this size. The stage 1 crowd was equivalent to a team of
1469.9 perfect classifiers; the stage 2 crowd was equivalent to a team of
102.4 perfect classifiers.

The total information generated during the survey is harder to interpret,
because it's exactly the high probability candidates that we leave in the
system that give rise to most of the information generated. What we can do is
divide the total information generated by the amount of information it takes
to classify a \sw subject all the way to the detection threshold (lens
probability 0.95), divide by this number, and then multiply by the survey
inefficiency to get a very rough estimate for the effective number of
detections corresponding to the crowd's contribution. With a prior probability
of $2\times10^{-4}$, it takes 11.6 bits to make a detection; the effective
number of detections contributed by the stage 1 and stage 2 volunteers is then
2670 and 24 respectively. These figures are close to the numbers of detections
given in column 7 of \Tref{tab:crowd:contributions}. 
The uncertainty in the interpretation of the
information generated provides further justification for our focus on the
expected information gain as a measure of volunteer contribution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}
\begin{center}
\caption{Total crowd and subject sample properties from the CFHTLS project.}
\label{tab:crowd:contributions}
\begin{tabular}{cccccccc}
  \hline
  \hline {Stage} & Subjects & Contribution                          & Agents & Skill      & Classifications          & Candidates & Information \\
                 & $\Ns$    & $\sum_k^{\Nv} \contribution_k$ (bits) & $\Nv$  & $\sum_k^{\Nv} \skill_k$ (bits) & $\sum_k^{\Nv} \thiseffort$ & $\Ncands$  & $\sum_j^{\Ns}\sum_k^{\Nv} \information_{j,k}$ (bits) \\
  \hline 
            1    & $427064$ & $1291997.4$ & $36929$ & $1469.9$ & $10768759$ & $3367$ & $91110.7$ \\
            2    & $3679$   &   $21895.8$ &  $1964$ &  $102.4$ &   $224745$ &   $91$ &  $1641.9$ \\
  \hline \hline
\end{tabular}
\medskip\\
\end{center}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
